US General Society Survey Analysis

### 2024-03-16 paper/paper.qmd
.

## Model 1
Initially, we filter the data to remove to exclude individuals who did not vote for Hillary Clinton or Donald Trump, due to the minimal votes for Gary Johnson, Jill Stein, Evan McMullin, and all other candidates which are insignificant to this paper. We first analyze the birth year of voters.

```{r}
#| echo: false
#| warning: false
#| message: false

# only choose values "Hillary Clinton" and "Donald Trump"
ces2020_votes <- ces2020 %>%
  filter(presvote16post != 'Did not vote') %>%
  filter(presvote16post %in% c("Hillary Clinton", "Donald Trump"))

# Calculate percentages
percentage_data <- ces2020_votes %>%
  group_by(birthyr) %>%
  summarise(HillaryClinton = sum(presvote16post == "Hillary Clinton") / n() * 100,
            DonaldTrump = sum(presvote16post == "Donald Trump") / n() * 100,
            Other = sum(presvote16post == "Other") / n() * 100,
            DidNotVote = sum(presvote16post == "Did not Vote") / n() * 100) %>%
  ungroup()

@@

plot_trump <- ggplot(percentage_data_long %>% filter(Candidate == "DonaldTrump"), aes(x = birthyr, y = Percentage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Percentage for Donald Trump", y = "Percentage (%)") +
  theme_minimal()
plot_trump

##porbably dont need this - cris
plot_other <- ggplot(percentage_data_long %>% filter(Candidate == "Other"), aes(x = birthyr, y = Percentage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Percentage for Other", y = "Percentage (%)") +
  theme_minimal()
plot_other

```
## Model 2
In this model, we conduct an examination of the relationship between voters' birth year and their gender for the 2016 and 2020 Presidential elections. This analysis is visualized by plotting a histogram that separate female voters on the left and male voters on the right, with voters' birth years measured along the x-axis, which ranges from 1925 to 2002. The y-axis quantifies the voter turnout for the year. For clarity ad symbolic representation, the colour blue was chosen to represent the Democratic candidates --Hillary Clinton for the 2016 election, and Joe Biden for the 2020 election, while red was chosen to the Represent the Republican candidate, Donald Trump, who sought the presidency in both terms. @fig-genderyr2016 present the distribution of votes in 2016, and @fig-genderyr2020 presents the data from the 2020 election.

>>>>

## Aging and Its Impact on Voting

In these following models we will of our research paper, we will look at the relationship between the demographic of age within the population and their voting patterns.The focus will primarily be on understanding how different age groups align themselves politically. By investigating the percentage of voters within each age bracket who support particular candidates, we aim to uncover nuanced insights into the demographic underpinnings of political support.Through the integration of descriptive statistics, analysis, and visual data representation we will offer a comprehensive overview on age base voting patterns. This will not only provide empirical evidence on the voting preferences of different ages but also help for a deeper understanding of the potential motivations behind their voting choices.

### Voters in 2016

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-hillary2016
#| fig-cap: "Linear Regression comparing Percentage of Votes For Hillary Clinton and Year of Birth (2016)"

# only choose values "Hillary Clinton" and "Donald Trump"
ces2020_votes <- ces2020 %>%
  filter(presvote16post %in% c("Hillary Clinton", "Donald Trump", "Other", "Did not vote"))

# Calculate percentages
percentage_data <- ces2020_votes %>%
  group_by(birthyr) %>%
  summarise(
    HillaryClinton = sum(presvote16post == "Hillary Clinton") / n() * 100,
    DonaldTrump = sum(presvote16post == "Donald Trump") / n() * 100,
    Other = sum(presvote16post == "Other") / n() * 100,
    DidNotVote = sum(presvote16post == "Did not vote") / n() * 100
  ) %>%
  ungroup()

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-trump2016
#| fig-cap: "Linear Regression comparing Percentage of Votes For Donald Trump and Year of Birth (2016)"

#Trum Plot
plot_trump <- ggplot(percentage_data_long %>% filter(Candidate == "DonaldTrump"), aes(x = birthyr, y = Percentage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Percentage for Donald Trump", y = "Percentage (%)") +
  theme_minimal()
plot_trump
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-other2016
#| fig-cap: "Linear Regression comparing Percentage of Votes For Other Caniddates and Year of Birth (2016)"

plot_other <- ggplot(percentage_data_long %>% filter(Candidate == "Other"), aes(x = birthyr, y = Percentage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "green") +
  labs(title = "Percentage for Other", x = "Birth Year", y = "Percentage (%)") +
  theme_minimal()
plot_other
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-dnv2016
#| fig-cap: "Linear Regression comparing Percentage of Non-Voters and Year of Birth (2016)"

#plot did not vote
plot_did_not_vote <- ggplot(percentage_data_long %>% filter(Candidate == "DidNotVote"), aes(x = birthyr, y = Percentage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "orange") +
  labs(title = "Percentage for Did Not Vote", x = "Birth Year", y = "Percentage (%)") +
  theme_minimal()
plot_did_not_vote
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-biden2020
#| fig-cap: "Linear Regression comparing Percentage of Votes For Joe Biden and Year of Birth (2020)"

ces2020_votes <- ces2020 %>%
  filter(voted_for %in% c("Biden", "Trump"))

# Calculate percentages for all categories
percentage_data_2 <- ces2020_votes %>%
  group_by(birthyr) %>%
  summarise(
    Biden = sum(voted_for == "Biden") / n() * 100,
    Trump = sum(voted_for == "Trump") / n() * 100
  ) %>%
  ungroup()

# Reshape data to long format
percentage_data_2_long <- pivot_longer(percentage_data_2, cols = -birthyr, names_to = "Candidate", values_to = "Percentage")

# Plot for Joe Biden
plot_biden <- ggplot(percentage_data_2_long %>% filter(Candidate == "Biden"), aes(x = birthyr, y = Percentage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Percentage for Joe Biden", y = "Percentage (%)", x = "Birth Year") +
  theme_minimal()

plot_biden

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-trump2020
#| fig-cap: "Linear Regression comparing Percentage of Votes For Donald Trump Year of Birth (2020)"

# Plot for Donald Trump
plot_trump_2 <- ggplot(percentage_data_2_long %>% filter(Candidate == "Trump"), aes(x = birthyr, y = Percentage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Percentage for Donald Trump", y = "Percentage (%)", x = "Birth Year") +
  theme_minimal()

plot_trump_2
```

## Model 2
In this model, we conduct an examination of the relationship between voters' birth year and their gender for the 2016 and 2020 Presidential elections. This analysis is visualized by plotting a histogram that separate female voters on the left and male voters on the right, with voters' birth years measured along the x-axis, which ranges from 1925 to 2002. The y-axis quantifies the voter turnout for the year. For clarity ad symbolic representation, the colour blue was chosen to represent the Democratic candidates --Hillary Clinton for the 2016 election, and Joe Biden for the 2020 election, while red was chosen to the Represent the Republican candidate, Donald Trump, who sought the presidency in both terms. @fig-genderyr2016 present the distribution of votes in 2016, and @fig-genderyr2020 presents the data from the 2020 election.
